Extended Data

### 2022-05-09 doc/submissions/2022_04_Nature/Response to reviews.Rmd

**We agree with this comment and specify our motivation in the methods section: "We defined CTW for *DOY~25~*, as opposed to other parameters describing the timing of growth, because spring temperatures should have the most direct influence on this variable."**

**We also repeat the analyses using a definition of CTW that is independent of *DOY~25~* and all other response variables. As explained in the methods, "To ensure that patterns were robust under an alternative definition of CTW, and to parallel the monthly time windows used in our tree-ring analysis (detailed below; Fig. 3, Extended Data Figure 6-7), we also ran analyses where we fixed the CTW identified by climwin to be the month with the most days in the CTW (Extended Data Table 2) each critical window."**

I appreciated the clarifications done concerning the data cleaning process. However, I didn’t find in the manuscript the quantity of data that have been removed at it stages. I believe it is worth mentioning it.

**We simplified the presentation of our data cleaning methods, ensuring that these numbers are clearly presented for both stagges of cleaning:**

- **Paragraph describing screening of raw dendrometer band data: "In total, this process removed 309 of the 2210 available records for 2011-2020 at SCBI and 1999-2003 at Harvard Forest."**
- **Paragraph describing removal of poorly fit records: "In total, this process removed 189 of the 2467 records deemed appropriate for analysis."**

### 2022-05-09 doc/manuscript/Dow_manuscript_Nature.Rmd
update text on n records removed, response to reviewer on this
#141

@camerondow35 , I've changed this to add direct statements of numbers removed, which I think is what the reviewer was looking for.

We removed tree-years with small or negligible total growth ($\Delta DBH$ $\le 0.005$ cm; SCBI = 26, Harvard Forest = 253), and tree-years where the first intra-annual measurement was later than the first spring survey (trees that were missed in the initial census; SCBI = 22, Harvard Forest = 8).
These were removed because they reduce the reliability of predicted growth in the modeled curves.

@@

After fitting the growth model, we removed tree-years with poor fits.
Models were judged to be poorly fit if modeled growth parameters were outliers, which were commonly indicative of unrealistic fits (e.g., growth occurring outside the growing season or over a very short period) and underlain by very slow tree growth or poor data records that passed the initial screening (described above).
Modeled fits for tree-years were removed under two conditions:
(1) *g~max~* was $\ge 2.5$ standard deviations away from the mean for each site-xylem architecture group combination (SCBI = 3, Harvard Forest = 11);
(2) timing variables (*DOY~ip~*, *DOY~25~*, *DOY~50~*, *DOY~75~*) were $\ge 2.5$ standard deviations away from the means for their site, xylem architecture group, and year (SCBI = 74, Harvard Forest = 101).
At both sites the tree-years removed through this method were proportional to the original sample size, indicating that no species or size class was disproportionately removed compared to others.
This process was repeated using 2 and 3 standard deviations as the cutoff for defining outliers, yielding qualitatively similar results.

>>>>

The raw dendrometer band data were screened to remove records or entire tree-years that were inappropriate for our analysis because of reduced reliability of predicted growth in the modeled curves.
Specifically, we removed tree-years with small or negligible total growth ($\Delta DBH$ $\le 0.005$ cm; SCBI = 26, Harvard Forest = 253) and tree-years where the first intra-annual measurement was later than the first spring survey (trees that were missed in the initial census; SCBI = 22, Harvard Forest = 8).
In total, this process removed `r 26+253+22+8` of the `r n_tree_years` available records for 2011-2020 at SCBI and 1999-2003 at Harvard Forest.

@@

After fitting the growth model, we removed tree-years with poor fits.
Models were judged to be poorly fit if modeled growth parameters were outliers, which were commonly indicative of unrealistic fits (e.g., growth occurring outside the growing season or over a very short period) and underlain by very slow tree growth or poor data records that passed the initial screening (described above).
Modeled fits for tree-years were removed under two conditions:
(1) *g~max~* was $\ge 2.5$ standard deviations away from the mean for each site-xylem architecture group combination (SCBI = 3, Harvard Forest = 11);
(2) timing variables (*DOY~ip~*, *DOY~25~*, *DOY~50~*, *DOY~75~*) were $\ge 2.5$ standard deviations away from the means for their site, xylem architecture group, and year (SCBI = 74, Harvard Forest = 101).
In total, this process removed `r 3+11+74+101` of the `r n_tree_years - 26+253+22+8` records deemed appropriate for analysis.
At both sites the tree-years removed through this method were proportional to the original sample size, indicating that no species or size class was disproportionately removed compared to others.
This process was repeated using 2 and 3 standard deviations as the cutoff for defining outliers, yielding qualitatively similar results.
